Equalizer\\Programming and User Guide

### 2012-06-26 Developer/ProgrammingGuide/paper.tex
Fix layout and included files

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/gpu-sd.pdf}
  {\caption{\label{fGPUSD}GPU discovery for auto-configuration}}
\end{wrapfigure}
Equalizer uses the GPU-SD library for discovering local and remote GPUs. Based
on this information, a typical configuration is dynamically compiled at
application start. The auto-configuration uses the same format as the static
configuration files, and can therefore be used to create template configurations
by running a test application and copying the generated configuration.

@@

Equalizer uses the library with the dns\_sd discovery module to gather the
information announced by all daemons on the local network, as well as the local
discovery modules for standalone auto-configuration. \fig{fGPUSD} shows how
GPU-SD is used to discover four remote GPUs on two different machines to use
them to render a single view on a laptop.

@@

If you are writing a configuration for workstation usage you can skip
the following sections and restart with \sref{sCompounds}.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/frusta.pdf}
  {\caption{\label{fFrusta}Wall and Projection Parameters}}
\end{wrapfigure}
A canvas has one or more layouts. One of the layouts is the active
layout, that is, this set of views is currently used for rendering. It
is possible to specify \textsf{OFF} as a layout, which deactivates the
canvas. It is possible to use the same layout on different canvases.

@@

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/layout.png}
  {\caption{\label{fLayout}Layout with four Views}}
\end{wrapfigure}
Enable the layout on each desired canvas by adding it to the
canvas. Since canvases reference layouts by name or index, layouts have
to be configured before their respective canvases in the configuration
file.

@@

\begin{figure}[ht!]\center
  \includegraphics[width=.9\textwidth]{images/wallLayout.jpg}
  {\caption{Display Wall using a six-Segment Canvas with a two-View Layout}}
\end{figure}

@@

A view has a fractional viewport relative to its layout.  A layout
is often fully covered by its views, but this is not a requirement.

@@

@@

\subsubsection{\label{sSwapBarrier}Synchronization - Swap Barriers}
Compounds may have a software or hardware swap barrier to synchronize the buffer
swap of multiple channels. Swap barriers are defined by the segment, which will
set the swap barrier on all its destination channel compounds.

A software swap barrier is configured by giving it a name. Windows with a swap
barrier of the same name synchronize with each other before executing the swap
buffer task. Before entering the barrier, \textsf{Window::finish} is called to
ensure that all OpenGL commands have been executed.  Swap barrier names are only
valid within the compound tree, that is, a compound from one compound tree
cannot be synchronized with a compound from another compound tree.

Display synchronization uses different algorithms. Framelock synchronizes each
vertical retrace of multiple graphic outputs using a hardware component. This is
configured in the driver, independently of the application. Genlock synchronizes
each horizontal and vertical retrace of multiple graphic outputs, but is not
used anymore for 3D graphics. Swap lock synchronizes the front and back buffer
swap of multiple windows, either using a software solution based on network
communication or a hardware solution based on a physical cable. It is
independent of, but often used in conjunction with framelock.

Framelock is used to synchronize the vertical retrace in multi-display active
stereo installations, e.g., for edge-blended projection systems or immersive
installations. It is combined with a software or hardware swap barrier. Software
barriers in this case can not guarantee that the buffer swap of all outputs
always happens with the same retrace. The time window for an output to miss the
retrace for the buffer swap is however very small, and the output will simply
swap on the next retrace, typically in 16 milliseconds.

@@

Equalizers are used to update compound parameters based on runtime
data. They are attached to a compound (sub-)tree, on which they
operate. The Equalizer distribution contains numerous example
configuration files using equalizers.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/lb.pdf}
  {\caption{\label{fLoadBalancing}\small 2D Load-Balancing}}
\end{wrapfigure}
Using a load equalizer is transparent to the application, and can be used with
any application for 2D, and with most applications for DB load-balancing. Some
applications do not support dynamic updates of the database range, and therefore
cannot be used with DB load-balancing.

@@

\begin{wrapfigure}{l}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/monitorEq.pdf}
  {\caption{\label{fMonitorEq}Monitoring a Projection Wall}}
\end{wrapfigure}
A monitor equalizer adjusts the frame zoom of the output frames used to
observe the rendering, depending on the destination channel size. The
output frames are downscaled on the GPU before readback, which results
in optimal performance.

@@

{\footnotesize\lstinputlisting[firstline=81, lastline=89]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=91, lastline=101]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=102, lastline=119]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=121, lastline=132]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=134, lastline=143]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=145, lastline=159]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=161, lastline=165]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=167, lastline=178]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=180, lastline=188]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=65, lastline=76]
  {../../../Equalizer/examples/eqPly/initData.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=123, lastline=130]
  {../../../Equalizer/examples/eqPly/frameData.h}}
{\footnotesize\lstinputlisting[firstline=55, lastline=84]
  {../../../Equalizer/examples/eqPly/frameData.cpp}}

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/modelDist.pdf}
  {\caption{\label{fModelDist}Scene Data in eqPly}}
\end{wrapfigure}

@@

\fig{fModelDist} shows the UML hierarchy of the model and distribution
classes. \sref{sNetObject2} illustrates other approaches to employ distributed
objects for data distribution and synchronization.

@@

{\footnotesize\lstinputlisting[firstline=142, lastline=174]
  {../../../Equalizer/examples/eqPly/vertexBufferDist.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=76, lastline=93]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=94, lastline=96]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=98, lastline=115]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=127, lastline=135]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=306, lastline=313]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=315, lastline=323]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=325, lastline=355]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

The event handling code is shortened here. A detailed description on how to
customize event handling can be found in \sref{sEventHandling}:

{\footnotesize\lstinputlisting[firstline=387, lastline=390]
  {../../../Equalizer/examples/eqPly/config.cpp}}
{\footnotesize\lstinputlisting[firstline=450, lastline=481]
  {../../../Equalizer/examples/eqPly/config.cpp}}
{\footnotesize\lstinputlisting[firstline=528, lastline=534]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=280, lastline=304]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=249, lastline=265]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

The object manager (OM) is a utility class and can be used to manage OpenGL
objects across shared contexts. Typically one OM is used for each set of shared
contexts of a single
GPU\footnote{\link{http://www.equalizergraphics.com/documents/design/objectManager.html}}.

@@

{\footnotesize\lstinputlisting[firstline=122, lastline=137]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=138, lastline=164]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=166, lastline=177]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=584, lastline=621]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

%\begin{wrapfigure}{r}{.618\textwidth}
\begin{figure}[h!t]\center
  \includegraphics[width=.9\textwidth]{images/render.pdf}
  {\caption{\label{fRender}Main Render Loop}}
\end{figure}
The model data is spatially organized in a 3-dimensional
kD-tree\footnote{See also \link{http://en.wikipedia.org/wiki/Kd-tree}}
for efficient view frustum culling. When the model is loaded by
\textsf{Node::config\-Init}, it is preprocessed into the kD-tree. During
this preprocessing step, each node of the tree gets a database range
assigned. The root node has the range [0, 1], its left child [0, 0.5]
and its right child [0.5, 1], and so on for all nodes in the tree. The
preprocessed model is saved in a binary format for accelerating
subsequent loading.

The rendering loop maintains a list of candidates to render, which
initially contains the root node. Each candidate of this list is tested
for full visibility against the frustum and range, and rendered if
visible. It is dropped if it is fully invisible or fully out of
range. If it is partially visible or partially in range, the children of
the node are added to the candidate list.

@@

Like most applications, eqPly uses most of the default implementation of the
\textsf{frameReadback} and \textsf{frameAssemble} task methods. To implement an
optimization and various customizations, \textsf{frameReadback} is
overwritten. eqPly does not need the alpha channel on the destination view. The
output frames are flagged to ignore alpha, which allows the compressor to drop
25\% of the data during image transfer. Furthermore, compression can be
disabled and the compression quality can be changed at runtime to demonstrate
the impact of compression on scalable rendering:

{\footnotesize\lstinputlisting[firstline=242, lastline=267]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \vspace{-2ex}\includegraphics[width=.382\textwidth]{images/eventFilter.pdf}
  {\caption{\label{fEventProcessing}Event Processing}}
\end{wrapfigure}
Events are received by an event handler. The event handler finds the
\textsf{eq::Sys\-tem\-Win\-dow} associated to the event. It then creates a
generic \textsf{Event}, which holds the event data in an independent format. The
original native event and this generic \textsf{Event} form the
\textsf{SystemWindowEvent}, which is passed to the concrete
\textsf{SystemWindow} for processing.

@@

\textsf{Window::processEvent} is responsible for handling
the event locally and for translating it into a generic
\textsf{ConfigEvent}. The config events are send to the application
thread using \textsf{Config::sendEvent}.

@@

{\footnotesize\lstinputlisting[firstline=36, lastline=44]
  {../../../Equalizer/examples/eVolve/error.h}}

@@

An implication of the draw\_sync and local\_sync models is that the application
node can not issue a new frame until it has completed all its draw tasks. For
larger cluster configurations it is therefore advisable to assign only assemble
operations to the application node. If the machine running the application
process should also contribute for rendering, a second node may be used to
perform off-screen draw and readback operations for the application node
process.

@@

{\footnotesize\lstinputlisting[firstline=138, lastline=154]
  {../../../Equalizer/libs/eq/client/agl/window.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=180, lastline=224]
  {../../../Equalizer/libs/eq/client/wgl/window.cpp}}

@@

The \textsf{declareRegion} is forwarded to \textsf{eq::Channel::declareRegion}.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/slices.pdf}
  {\caption{\label{fSlices}Volume Rendering using 3D Textures}}
\end{wrapfigure}
When using 2D (sort-first) or stereo decompositions, no special
programming is needed to achieve good scalability, as \textsf{eVolve} is
mostly fill-limited and therefore scales nicely in these modes.

@@

Conceptually, the individual volume bricks of each of the source
channels produces pixel data which can be handled like one big 'slice'
through the full texture. Therefore they have to be blen\-ded
back-to-front in the same way as the slice planes are blended during
rendering.

@@

The supersampling algorithm used by eqPly performs anti-aliasing, using a
technique which computes samples within a pixel by jittering the frustum by a
subpixel amount. The pixel is split into an 16x16 subpixel grid, and a sample is
taken randomly within each sub-pixel. The samples are accumulated and displayed
after each step.

@@

{\footnotesize\lstinputlisting[firstline=460, lastline=483]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=320, lastline=344]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=542, lastline=563]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=491, lastline=497]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=503, lastline=529]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=176, lastline=182]
  {../../../equalizergraphics/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=184, lastline=187]
  {../../../equalizergraphics/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=211, lastline=236]
  {../../../equalizergraphics/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=238, lastline=259]
  {../../../equalizergraphics/documents/design/fileFormat.shtml}}

>>>>

Equalizer uses the GPU-SD library for discovering local and remote GPUs. Based
on this information, a typical configuration is dynamically compiled during the
initialization of the application. The auto-configuration uses the same format
as the static configuration files. It can be used to create template
configurations by running a test application or the eqServer binary and using
the generated configuration, saved as \textless session\textgreater.auto.eqc.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/gpu-sd.pdf}
  {\caption{\label{fGPUSD}GPU discovery for auto-configuration}}
\end{wrapfigure}
Equalizer uses the library with the dns\_sd discovery module to gather the
information announced by all daemons on the local network, as well as the local
discovery modules for standalone auto-configuration. The default configuration
is using all locally discovered GPUs.

\fig{fGPUSD} shows how GPU-SD is used to discover four remote GPUs on two
different machines to use them to render a single view on a laptop.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/frusta.pdf}
  {\caption{\label{fFrusta}Wall and Projection Parameters}}
\end{wrapfigure}
If you are writing a configuration for workstation usage you can skip
the following sections and restart with \sref{sCompounds}.

@@

A canvas has one or more layouts. One of the layouts is the active
layout, that is, this set of views is currently used for rendering. It
is possible to specify \textsf{OFF} as a layout, which deactivates the
canvas. It is possible to use the same layout on different canvases.

@@

\subsubsection{\label{sSwapBarrier}Swap and Frame Synchronization}
Canvases and segments may have a software or hardware swap barrier to
synchronize the buffer swap of multiple channels. The swap barriers are
inherited from the canvas to the segment ultimately to all compounds using a
destination channel of the segment.

A software swap barrier is configured by giving it a name. Windows with a swap
barrier of the same name synchronize with each other before executing the swap
buffer task. Before entering the barrier, \textsf{Window::finish} is called to
ensure that all OpenGL commands have been executed.  Swap barrier names are only
valid within the compound tree, that is, a compound from one compound tree
cannot be synchronized with a compound from another ompound tree.

Display synchronization uses different algorithms. Framelock synchronizes each
vertical retrace of multiple graphic outputs using a hardware component. This is
configured in the driver, independently of the application. Genlock synchronizes
each horizontal and vertical retrace of multiple graphic outputs, but is not
commonly used anymore for 3D graphics. Swap lock synchronizes the front and back
buffer swap of multiple windows, either using a software solution based on
network communication or a hardware solution based on a physical cable. It is
independent of, but often used in conjunction with framelock.

Framelock is used to synchronize the vertical retrace in multi-display active
stereo installations, e.g., for edge-blended projection systems or immersive
installations. It is combined with a software or hardware swap barrier. Software
barriers in this case cannot guarantee that the buffer swap of all outputs
always happens with the same retrace. The time window for an output to miss the
retrace for the buffer swap is however very small, and the output will simply
swap on the next retrace, typically in 16 milliseconds.

@@

Enable the layout on each desired canvas by adding it to the
canvas. Since canvases reference layouts by name or index, layouts have
to be configured before their respective canvases in the configuration
file.

@@

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/layout.png}
  {\caption{\label{fLayout}Layout with four Views}}
\end{wrapfigure}
A view has a fractional viewport relative to its layout.  A layout
is often fully covered by its views, but this is not a requirement.

@@

\fig{fLayout} shows an example layout using four views on a single
segment. \fig{fDisplay} shows a real-world setup of a single canvas with six
segments using underlap, with a two-view layout activated. This configuration
generates eight destination channels.

\begin{figure}[ht!]\center
  \includegraphics[width=.9\textwidth]{images/wallLayout.jpg}
  {\label{fDisplay}\caption{Display Wall using a six-Segment Canvas with a two-View Layout}}
\end{figure}

@@

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/lb.pdf}
  {\caption{\label{fLoadBalancing}\small 2D Load-Balancing}}
\end{wrapfigure}
Equalizers are used to update compound parameters based on runtime
data. They are attached to a compound (sub-)tree, on which they
operate. The Equalizer distribution contains numerous example
configuration files using equalizers.

@@

Using a load equalizer is transparent to the application, and can be used with
any application for 2D, and with most applications for DB load-balancing. Some
applications do not support dynamic updates of the database range, and therefore
cannot be used with DB load-balancing.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/monitorEq.pdf}
  {\caption{\label{fMonitorEq}Monitoring a Projection Wall}}
\end{wrapfigure}
A monitor equalizer adjusts the frame zoom of the output frames used to
observe the rendering, depending on the destination channel size. The
output frames are downscaled on the GPU before readback, which results
in optimal performance.

@@

{\footnotesize\lstinputlisting[firstline=83, lastline=91]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=93, lastline=103]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=104, lastline=121]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=123, lastline=134]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=136, lastline=145]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=147, lastline=161]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=163, lastline=167]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=169, lastline=180]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=182, lastline=190]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=66, lastline=77]
  {../../../Equalizer/examples/eqPly/initData.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=123, lastline=130]
  {../../../Equalizer/examples/eqPly/frameData.h}}
{\footnotesize\lstinputlisting[firstline=54, lastline=82]
  {../../../Equalizer/examples/eqPly/frameData.cpp}}

@@

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/modelDist.pdf}
  {\caption{\label{fModelDist}Scene Data in eqPly}}
\end{wrapfigure}
\fig{fModelDist} shows the UML hierarchy of the model and distribution
classes. \sref{sNetObject2} illustrates other approaches to employ distributed
objects for data distribution and synchronization.

@@

{\footnotesize\lstinputlisting[firstline=144, lastline=176]
  {../../../Equalizer/examples/eqPly/vertexBufferDist.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=69, lastline=86]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=92, lastline=93]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=96, lastline=113]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=125, lastline=133]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=294, lastline=301]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=303, lastline=311]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=313, lastline=343]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

The event handling code repdroduced here is just showing the handling of one
type of event. A detailed description on how to customize event handling can be
found in \sref{sEventHandling}:

{\footnotesize\lstinputlisting[firstline=471, lastline=476]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=265, lastline=292]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=247, lastline=263]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

The object manager (OM) is a utility class and can be used to manage OpenGL
objects across shared contexts. Typically one OM is used for each set of shared
contexts of a single GPU.

@@

{\footnotesize\lstinputlisting[firstline=124, lastline=144]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=146, lastline=172]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=174, lastline=185]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=593, lastline=626]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

The model data is spatially organized in a 3-dimensional
kD-tree\footnote{See also \link{http://en.wikipedia.org/wiki/Kd-tree}}
for efficient view frustum culling. When the model is loaded by
\textsf{Node::config\-Init}, it is preprocessed into the kD-tree. During
this preprocessing step, each node of the tree gets a database range
assigned. The root node has the range [0, 1], its left child [0, 0.5]
and its right child [0.5, 1], and so on for all nodes in the tree. The
preprocessed model is saved in a binary format for accelerating
subsequent loading.

\begin{wrapfigure}{r}{.618\textwidth}
%\begin{figure}[h!t]\center
  \includegraphics[width=.618\textwidth]{images/render.pdf}
  {\caption{\label{fRender}Main Render Loop}}
\end{wrapfigure}
%\end{figure}
The rendering loop maintains a list of candidates to render, which
initially contains the root node. Each candidate of this list is tested
for full visibility against the frustum and range, and rendered if
visible. It is dropped if it is fully invisible or fully out of
range. If it is partially visible or partially in range, the children of
the node are added to the candidate list.

@@

Like most applications, eqPly uses most of the default implementation of the
\textsf{frameReadback} and \textsf{frameAssemble} task methods. To implement an
optimization and various customizations, \textsf{frame\-Read\-back} is
overwritten. eqPly does not need the alpha channel on the destination view. The
output frames are flagged to ignore alpha, which allows the compressor to drop
25\% of the data during image transfer. Furthermore, compression can be
disabled and the compression quality can be changed at runtime to demonstrate
the impact of compression on scalable rendering:

{\footnotesize\lstinputlisting[firstline=256, lastline=275]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

@@

Events are received by an event handler. The event handler finds the
\textsf{eq::Sys\-tem\-Win\-dow} associated to the event. It then creates a
generic \textsf{Event}, which holds the event data in an independent format. The
original native event and this generic \textsf{Event} form the
\textsf{SystemWindowEvent}, which is passed to the concrete
\textsf{SystemWindow} for processing.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \vspace{-2ex}\includegraphics[width=.382\textwidth]{images/eventFilter.pdf}
  {\caption{\label{fEventProcessing}Event Processing}}
\end{wrapfigure}
\textsf{Window::processEvent} is responsible for handling
the event locally and for translating it into a generic
\textsf{ConfigEvent}. The config events are send to the application
thread using \textsf{Config::sendEvent}.

@@

{\footnotesize\lstinputlisting[firstline=36, lastline=45]
  {../../../Equalizer/examples/eVolve/error.h}}

@@

An implication of the draw\_sync and local\_sync models is that the application
node cannot issue a new frame until it has completed all its draw tasks. For
larger cluster configurations it is therefore advisable to assign only assemble
operations to the application node. If the machine running the application
process should also contribute for rendering, a second node may be used to
perform off-screen draw and readback operations for the application node
process.

@@

{\footnotesize\lstinputlisting[firstline=142, lastline=158]
  {../../../Equalizer/libs/eq/client/agl/window.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=185, lastline=229]
  {../../../Equalizer/libs/eq/client/wgl/window.cpp}}

@@

The \textsf{declareRegion} is eventually forwarded to
\textsf{eq::Channel::declareRegion}.

@@

When using 2D (sort-first) or stereo decompositions, no special
programming is needed to achieve good scalability, as \textsf{eVolve} is
mostly fill-limited and therefore scales nicely in these modes.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/slices.pdf}
  {\caption{\label{fSlices}Volume Rendering using 3D Textures}}
\end{wrapfigure}
Conceptually, the individual volume bricks of each of the source
channels produces pixel data which can be handled like one big 'slice'
through the full texture. Therefore they have to be blen\-ded
back-to-front in the same way as the slice planes are blended during
rendering.

@@

The supersampling algorithm used by eqPly performs anti-aliasing, using a
technique which computes randomized samples within a pixel by jittering the
frustum by a subpixel amount. The pixel is split into an 16x16 subpixel grid,
and a sample is taken randomly within each sub-pixel. The samples are
accumulated and displayed after each step.

@@

{\footnotesize\lstinputlisting[firstline=469, lastline=493]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=329, lastline=353]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=551, lastline=572]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=500, lastline=506]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=512, lastline=538]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=177, lastline=183]
  {../../../equalizergraphics/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=185, lastline=188]
  {../../../equalizergraphics/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=212, lastline=237]
  {../../../equalizergraphics/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=239, lastline=260]
  {../../../equalizergraphics/documents/design/fileFormat.shtml}}
